---
title: "Beads"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
library(cowplot)

library(ggrepel)
library(pbapply)
library(scales)

theme_set(theme_cowplot())

dir.create("../figures/main_figure_2/", showWarnings = FALSE)
dir.create("../figures/sup_figure_2/", showWarnings = FALSE)

source("../scripts/notebooks_helpers/beads_and_tso_only.R")
source("../scripts/helpers/scales.R")

knitr::opts_chunk$set(echo = TRUE)
```

```{r}
rrn1 <- c(15062083, 15063836)
rrn2 <- c(15064301, 15064453)
rrn3 <- c(15064838, 15068346)
```

## Load data

```{r}
geneid <- fread("../data/annotations/c_elegans.PRJNA13758.WS265.geneIDs.txt.gz")
geneid[V3 == "", V3 := V4]
geneid <- geneid[, .(GeneName = V2, GeneSymbol = V3, GeneType = V6)]
geneid <- geneid[! duplicated(GeneName)]
```

```{r}
beads <- readRDS("../data/formated_counts/counts_and_annots_beads.rds")
beads[["da"]] <- fread("../data/differential_analysis/tables/beads/beads.csv.gz")
beads[["norm"]] <- sweep(beads$counts, 2, beads$nf, "/")
beads$da <- merge(beads$da, geneid[, .(GeneName, GeneType)])

tso <- readRDS("../data/formated_counts/counts_and_annots_tso_only.rds")
tso[["da"]] <- fread("../data/differential_analysis/tables/tso_only/tso_only.csv.gz")
tso[["norm"]] <- sweep(tso$counts, 2, tso$nf, "/")
tso$da <- merge(tso$da, geneid[, .(GeneName, GeneType)])
```

```{r}
beads_genomecov <- loadGenomeCov(beads$annots)
beads_genomecov[, NF := beads$nf[Sample]]

tso_genomecov <- loadGenomeCov(tso$annots)
tso_genomecov[, NF := tso$nf[Sample]]
```

## Genome Coverage

```{r}
gc <- tso_genomecov[, .(Coverage = mean(Count/NF)), by = c("Position", "TSOOnly")]

gp_cov_std <- plotCov(gc[TSOOnly == FALSE &
                           
                           Position >= rrn1[1]-500 & 
                           Position <= rrn3[2]+500]) +
  ggtitle("Standard (oligo-dT and TSO)") +
  theme(plot.title = element_text(size = 15, face = "bold"))

gp_cov_tso <- plotCov(gc[TSOOnly == TRUE &
                           Position >= rrn1[1]-500 & 
                           Position <= rrn3[2]+500]) +
  ggtitle("TSO Only") +
  theme(plot.title = element_text(size = 15, face = "bold"))

gp_cov_std
ggsave("../figures/sup_figure_2/beads_and_tso_only_genomecov_standard.pdf", width = 7, height = 6)

gp_cov_tso
ggsave("../figures/sup_figure_2/beads_and_tso_only_genomecov_tso.pdf", width = 7, height = 6)
```
## ECDF of raw gene counts showing reduction of rrn-3.1 in the library

```{r}
set.seed(63)
std <- beads$counts[, beads$annots[Beads == FALSE]$Sample]
bd <- beads$counts[, beads$annots[Beads == TRUE]$Sample]
prop_df <- data.table(GeneName = rownames(std), 
                      Stantard =  rowMeans(sweep(std, 2, colSums(std), "/")),
                      Beads = rowMeans(sweep(bd, 2, colSums(bd), "/")))
rm(std, bd)
prop_df <- melt(prop_df, id.vars = "GeneName", value.name = "Proportion", variable.name = "Condition")
prop_df <- prop_df[Proportion != 0]
prop_df <- merge(geneid, prop_df, by = "GeneName", all.x = FALSE, all.y = TRUE)

prop_df <- prop_df[order(Proportion, decreasing = TRUE)]
prop_df <- prop_df[, Number := 1:.N, by = .(Condition)]
prop_df <- prop_df[, CDF := 1 - Number/.N, by = .(Condition)]

prop_df[Number <= 5, Label := paste0(GeneSymbol, " (", round(100*Proportion, 1), "%)")]
prop_df[Number <= 5, Label := factor(Label, levels = unique(Label))]

gp_prop_beads <- ggplot(prop_df, aes((Proportion), CDF, color = Condition)) +
  geom_point() +
  scale_x_log10(labels = function(x) {   parse(text=gsub("e\\+*", " %*% 10^", scales::scientific_format()(x))) }) + 
  scale_color_manual(name = "Protocol", 
                     values = c("gray0", "orangered"),
                     labels = c("Standard", "Beads")) +
  geom_text_repel(aes(label = Label), show.legend = FALSE,
                  size = 3.7,
                  prop_df[Number <= 3 & Condition == "Beads"],
                  box.padding = 0.45,
                  arrow = arrow(length = unit(0.04, "npc")),
                  direction    = "x",
                  nudge_x      = -1,
                  nudge_y      = -1,
                  angle        = 0,
                  vjust        = 0,
                  hjust        = 1,
                  segment.size = 0.5) +
  ylab("Empirical Cumulative Probability") +
  xlab("Proportion of Gene in Read Library") +
  theme(legend.position = "top")

gp_prop_beads
ggsave("../figures/main_figure_2/beads_and_tso_only_cdf_beads.pdf", width = 7, height = 6)
```

## Differential analysis comparing TSO-only with the standard protocol. 

```{r}
ggplot(tso$da[! is.na(GeneType)], 
       aes(TSOOnly_log2FoldChange, 
           TSOOnly_padj,
           color = GeneType == "protein_coding_gene")) +
  geom_point() +
  scale_y_continuous(trans=reverselog_trans(10), labels = scientific_10) +
  xlab("log2 Fold Change") +
  ylab("Adjusted p-value") +
  geom_hline(yintercept = 1) +
  scale_color_manual(name = "", 
                     values = c("#990000ff", "#0b5394ff"),
                     labels = c("Non protein coding genes", 
                                "Protein coding genes")) +
  theme(legend.position = "top")
```

## Differential analysis comparing cDNA generated on oligo-dT carrying magnetic beads with the standard protocol. 

```{r}
set.seed(65)
gp_volcano_beads <- ggplot(beads$da, 
                           aes(Beads_log2FoldChange, 
                               Beads_padj,
                               color = Beads_padj < 0.001)) +
  geom_point() +
  scale_y_continuous(trans=reverselog_trans(10), labels = scientific_10) +
  xlab("log2 Fold Change\nBeads vs Standard") +
  ylab("Adjusted p-value") +
    scale_color_manual(name = "Effect of Beads",
                     values = c("gray0", "orangered"),
                     labels = c("p-value > 0.001",
                                "p-value < 0.001")) +
  geom_text_repel(aes(label = GeneSymbol), 
                  beads$da[(Beads_padj < 1e-15 & 2**Beads_log2FoldChange < 0.2) | (Beads_log2FoldChange < -6 & Beads_padj < 1e-4)],
                  min.segment.length = unit(0, 'lines'),
                  show.legend = FALSE) +
  theme(legend.position = "top")
gp_volcano_beads
```

## Comparing the effect of removing the oligo-dT in the retro-transcriptionwith the effect of using oligo-dT carrying magnetic beads.

```{r}
da <- merge(tso$da, beads$da, by = "GeneName")
da[, BeadsSignificant := Beads_padj < 1e-3]

gp_compare <- ggplot(da[order(BeadsSignificant)],
                     aes(TSOOnly_log2FoldChange,
                         Beads_log2FoldChange,)) +
  geom_point(aes(color = BeadsSignificant)) +
  scale_color_manual(name = "Effect of Beads",
                     values = c("gray0", "orangered"),
                     labels = c("p-value > 0.001",
                                "p-value < 0.001")) +
  theme(legend.position = "top") +
  xlab("log2 Fold Change\nTSO Only vs Standard") +
  ylab("log2 Fold Change\nBeads vs Standard")
gp_compare
```
