---
title: "PCA"
output: html_document
---

```{r setup, include=FALSE}
library(data.table)
library(ggplot2)
library(cowplot)
library(Rtsne)

dir.create("../figures/main_figure_2/", showWarnings = FALSE)
dir.create("../figures/sup_figure_2/", showWarnings = FALSE)

theme_set(theme_cowplot())

source("../scripts/helpers/preprocess.R")

knitr::opts_chunk$set(echo = TRUE)
```

## Load data

```{r}
genotype_colors <- c(N2 = "#990000ff", `daf-2(e1368)` = "#0b5394ff")
```


```{r}
data <- readRDS("../data/formated_counts/counts_and_annots_single_and_pooled_worms.rds")
bc <- readRDS("../data/batch_corrected_counts/single_and_pooled_worms.rds")
data <- c(data, setNames(bc$batch_counts_list, paste0("counts_", names(bc$batch_counts_list))))
rm(bc)
```

```{r}
counts_list <- data[grepl("counts", names(data))]
counts_list <- counts_list[! sapply(counts_list, is.null)]

norm_list <- lapply(counts_list, function(x) {
  # filterGenes(x, data$annots[colnames(x)]$Group)
  k <- normalizationFactor(x, groups = data$annots[colnames(x)]$Group)
  normalizeCounts(x, k)
})

pca_list <- lapply(norm_list, function(norm) {
  norm <- filterGenes(norm, min_count = 5, min_prop = 0.8)
  pca <- prcomp(log(t(norm+1)), scale = TRUE, center = TRUE)
  pca_df <- cbind(data$annots[colnames(norm), on = "Sample"], pca$x[, 1:2])
  pca_df
})
```

```{r}
pw_da <- fread("../data/differential_analysis/tables/single_and_pooled_worms/d8_pw.csv.gz")
sw_da <- fread("../data/differential_analysis/tables/single_and_pooled_worms/d8_sw.csv.gz")
```

## Compare mean pooled and single

```{r}
sw_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & NumberOfWorms == 1 & Genotype == "N2"]$Sample])
pw_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & NumberOfWorms != 1 & Genotype == "N2"]$Sample])

plot(sw_mu+1, pw_mu+1, log="xy", size = .5)
abline(b=1, a=0)
```

```{r}
sw_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & NumberOfWorms == 1 & Genotype == "daf-2(e1368)"]$Sample])
pw_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & NumberOfWorms != 1 & Genotype == "daf-2(e1368)"]$Sample])

plot(sw_mu+1, pw_mu+1, log="xy")
abline(b=1, a=0)
```

```{r}
br1_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & BiologicalReplicate == 1 & Genotype == "N2"]$Sample])
br5_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & BiologicalReplicate == 5 & Genotype == "N2"]$Sample])

plot(br1_mu+1, br5_mu+1, log="xy")
abline(b=1, a=0)
```

```{r}
br1_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & BiologicalReplicate == 1 & Genotype == "daf-2(e1368)"]$Sample])
br5_mu <- rowMeans(norm_list$counts_d8_aw[, data$annots[Day == 8 & BiologicalReplicate == 5 & Genotype == "daf-2(e1368)"]$Sample])

plot(br1_mu+1, br5_mu+1, log="xy")
abline(b=1, a=0)
```

## All conditions

### Day 1 N2

```{r}
br1_mu <- rowMeans(norm_list$counts_bd_sw[, data$annots[Day == 1 & BiologicalReplicate == 1 &
                                                          Genotype == "N2" & NumberOfWorms == 1]$Sample])
br5_mu <- rowMeans(norm_list$counts_bd_sw[, data$annots[Day == 1 & BiologicalReplicate == 5 & 
                                                          Genotype == "N2" & NumberOfWorms == 1]$Sample])

plot(br1_mu+1, br5_mu+1, log="xy")
abline(b=1, a=0)
```
### Day 8 N2

```{r}
br1_mu <- rowMeans(norm_list$counts_bd_sw[, data$annots[Day == 8 & BiologicalReplicate == 1 & 
                                                          Genotype == "N2" & NumberOfWorms == 1]$Sample])
br5_mu <- rowMeans(norm_list$counts_bd_sw[, data$annots[Day == 8 & BiologicalReplicate == 5 & 
                                                          Genotype == "N2" & NumberOfWorms == 1]$Sample])

plot(br1_mu+1, br5_mu+1, log="xy")
abline(b=1, a=0)
```

### Day 8 daf-2

```{r}
br1_mu <- rowMeans(norm_list$counts_bd_sw[, data$annots[Day == 8 & BiologicalReplicate == 1 & 
                                                          Genotype == "daf-2(e1368)" & NumberOfWorms == 1]$Sample])
br5_mu <- rowMeans(norm_list$counts_bd_sw[, data$annots[Day == 8 & BiologicalReplicate == 5 & 
                                                          Genotype == "daf-2(e1368)" & NumberOfWorms == 1]$Sample])

plot(br1_mu+1, br5_mu+1, log="xy")
abline(b=1, a=0)
```

## Do all PCAs to check if batch correction is working

```{r, fig.width=9, fig.height=8}
for (i in names(pca_list)) {
  pca_df <- copy(pca_list[[i]])
  pca_df[, SingleWorm := NumberOfWorms == 1]
  gp <- ggplot(pca_df, aes(PC1, PC2, color = BiologicalReplicate, shape = SingleWorm)) +
    geom_point() +
    facet_wrap(paste("Day", Day) ~ Genotype, scales = "free") +
    ggtitle(i) +
    theme(legend.position = "top") +
    scale_shape_manual(values = c(16, 1))
  plot(gp)
}
```

## Single worm and pooled worm PCA

```{r}
nosel_counts <- counts_list$counts_d8_aw
annots <- data$annots[colnames(nosel_counts)]
nf <- normalizationFactor(nosel_counts, groups = annots$Group)

sel_counts <- nosel_counts[sw_da[Strain_QZ120_vs_QZ0_padj < 1e-2]$GeneName, annots$Sample]

nosel_norm <- normalizeCounts(nosel_counts, nf)
sel_norm <- normalizeCounts(sel_counts, nf)

nosel_lognorm <- scale(t(log(nosel_norm+1)))
sel_lognorm <- scale(t(log(sel_norm+1)))

pw_samples <- annots[NumberOfWorms != 1]$Sample
sw_samples <- annots[NumberOfWorms == 1]$Sample
```

### Select genes

#### Run PCA on pooled worms, and project single worms

```{r}
pca <- prcomp(sel_lognorm[pw_samples, ])
pca_df <- cbind(annots, sel_lognorm %*% pca$rotation)
varexp <- matrixStats::colVars(pca$x)

ggplot(pca_df, aes(PC1, PC2, color = Genotype, size = NumberOfWorms == 1)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  xlab(paste0("Principal Component 1 (", round(100 * varexp[1]/sum(varexp), 1), "%)")) +
  ylab(paste0("Principal Component 2 (", round(100 * varexp[2]/sum(varexp), 2), "%)"))
ggsave("../figures/main_figure_2/pca_single_worm_day8_projection_from_pooled_worms.pdf", width = 7, height = 6)
```

#### Run PCA on single worms, and project pooled worms

```{r}
pca <- prcomp(sel_lognorm[sw_samples, ])
pca_df <- cbind(annots, sel_lognorm %*% pca$rotation)
varexp <- matrixStats::colVars(pca$x)

ggplot(pca_df, aes(PC1, PC2, color = Genotype, size = NumberOfWorms == 1)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  xlab(paste0("Principal Component 1 (", round(100 * varexp[1]/sum(varexp), 1), "%)")) +
  ylab(paste0("Principal Component 2 (", round(100 * varexp[2]/sum(varexp), 2), "%)"))
ggsave("../figures/main_figure_2/pca_single_worm_day8_projection_from_single_worms.pdf", width = 7, height = 6)
```

#### Run PCA on all worms

```{r}
pca <- prcomp(sel_lognorm)
pca_df <- cbind(annots, pca$x[, 1:2])
varexp <- pca$sdev^2

ggplot(pca_df, aes(PC1, PC2, color = Genotype, size = NumberOfWorms == 1)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  xlab(paste0("Principal Component 1 (", round(100 * varexp[1]/sum(varexp), 1), "%)")) +
  ylab(paste0("Principal Component 2 (", round(100 * varexp[2]/sum(varexp), 2), "%)"))
ggsave("../figures/sup_figure_2/pca_single_worm_day8_all_samples.pdf", width = 7, height = 6)
```

#### Run t-SNE on all worms

```{r}
tsne_out <- Rtsne(sel_lognorm, k = 2, perplexity = 50, theta = 0, initial_dims = 50, max_iter = 10000)
tsne_df <- cbind(annots, tsne_out$Y[, 1:2])

ggplot(tsne_df, aes(V1, V2, color = Genotype, size = NumberOfWorms == 1)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  xlab(paste0("t-SNE Dimension 1")) +
  ylab(paste0("t-SNE Dimension 2"))
ggsave("../figures/sup_figure_2/tsne_single_worm_day8_all_samples.pdf", width = 7, height = 6)
```

### No selection of genes

#### Run PCA on pooled worms, and project single worms

```{r}
pca <- prcomp(nosel_lognorm[pw_samples, ])
pca_df <- cbind(annots, nosel_lognorm %*% pca$rotation)
varexp <- matrixStats::colVars(pca$x)

ggplot(pca_df, aes(PC1, PC2, color = Genotype, size = NumberOfWorms == 1)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  xlab(paste0("Principal Component 1 (", round(100 * varexp[1]/sum(varexp), 1), "%)")) +
  ylab(paste0("Principal Component 2 (", round(100 * varexp[2]/sum(varexp), 2), "%)"))
ggsave("../figures/sup_figure_2/pca_single_worm_day8_projection_from_pooled_worms_all_genes.pdf", width = 7, height = 6)
```

#### Run PCA on single worms, and project pooled worms

```{r}
pca <- prcomp(nosel_lognorm[sw_samples, ])
pca_df <- cbind(annots, nosel_lognorm %*% pca$rotation)
pca_df[, SingleWorm := NumberOfWorms == 1]
varexp <- matrixStats::colVars(pca$x)

ggplot(pca_df, aes(PC1, PC2, color = Genotype, size =SingleWorm)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  geom_point(data = pca_df[SingleWorm == TRUE, .(PC1 = mean(PC1), PC2 = mean(PC2)), by = .(Genotype)], pch = 1, size = 4) + 
  xlab(paste0("Principal Component 1 (", round(100 * varexp[1]/sum(varexp), 1), "%)")) +
  ylab(paste0("Principal Component 2 (", round(100 * varexp[2]/sum(varexp), 2), "%)"))
ggsave("../figures/sup_figure_2/pca_single_worm_day8_projection_from_single_worms_all_genes_with_mean.pdf", width = 7, height = 6)
```

#### Run PCA on all worms

```{r}
pca <- prcomp(nosel_lognorm)
pca_df <- cbind(annots, pca$x[, 1:2])
varexp <- pca$sdev^2

ggplot(pca_df, aes(PC1, PC2, color = Genotype, size = NumberOfWorms == 1)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  xlab(paste0("Principal Component 1 (", round(100 * varexp[1]/sum(varexp), 1), "%)")) +
  ylab(paste0("Principal Component 2 (", round(100 * varexp[2]/sum(varexp), 2), "%)"))
ggsave("../figures/sup_figure_2/pca_single_worm_day8_all_samples_all_genes.pdf", width = 7, height = 6)
```

#### Run t-SNE on all worms

```{r}
tsne_out <- Rtsne(nosel_lognorm, k = 2, perplexity = 50, theta = 0, initial_dims = 50, max_iter = 10000)
tsne_df <- cbind(annots, tsne_out$Y[, 1:2])

ggplot(tsne_df, aes(V1, V2, color = Genotype, size = NumberOfWorms == 1)) +
  geom_point() +
  scale_color_manual(values = genotype_colors) +
  scale_size_manual(values = c(1.7, .5), name = "Number of Worms", labels = c("Pooled", "Single")) + 
  xlab(paste0("t-SNE Dimension 1")) +
  ylab(paste0("t-SNE Dimension 2"))
ggsave("../figures/sup_figure_2/tsne_single_worm_day8_all_samples_all_genes.pdf", width = 7, height = 6)
```
